Supporting information: Copulas and their potential for ecology

### 2019-09-21 SupportingInformation.Rmd
edits missed before

The data of Fig. \ref{MT-fig_cop_pedag1}A-B were generated to have standard normal marginals.
Fig. \ref{fig_transf_pearson_spearman}A can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
of both variables by the cummulative distribution function (cdf) of the standard normal distribution, and then by the inverse of 
the cdf of the gamma distribution with shape and scale parameters 2. 
The functions `pnorm` and `qgamma` in R were used.
Fig. \ref{fig_transf_pearson_spearman}B can be obtained from Fig. \ref{MT-fig_cop_pedag1}B in the same way.
The Pearson correlations differ substantially after transformation, 
but the Spearman correlations are unchanged (Figs \ref{MT-fig_cop_pedag1}A-B, \ref{fig_transf_pearson_spearman}).
A scatterploy very simlar to Fig. \ref{MT-fig_cop_pedag1}C can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
of both variables by the cdf of the standard normal distribution.
Likewise a scatterplot similar to Fig. \ref{MT-fig_cop_pedag1}D can be obtained from Fig. \ref{MT-fig_cop_pedag1}B in the same way.

@@

Now let $D$ be any bivariate copula and let 
$G_A$ and $G_B$ be univariate distribution functions, so that 
$D(G_A(x),G_B(y))$ is guaranteed by Sklar's theorem to be a a distribution function.
We here show that if $G_A$ and $G_B$ and continuous and strictly monotonic, then
$D(G_A(x),G_B(y))$ is the distribution function of the random variable 
$(G_A^{-1}(U),G_B^{-1}(V))$, where $(U,V)$ now denotes the random
vector with uniform marginals associated with $D$ and
$G_A^{-1}$ and $G_B^{-1}$ are the inverses of $G_A$ and $G_B$.
The inverses exist because of the strict monotonicity assumption.
To get to the desired result, note that the distribution function of 
$(G_A^{-1}(U),G_B^{-1}(V))$ is $P[G_A^{-1}(U) \leq x, G_B^{-1}(V) \leq y]$,
which equals $P[U \leq G_A(x), V \leq G_B(y)]$, by strict monotonicity.
But this is $D(G_A(x),G_B(y))$ because $D$ is the distribution function of 
$(U,V)$. In fact, a stronger statement assuming
neither continuity nor strict monotonicity is possible, if one uses generalized
inverses for $G_A^{-1}$ and $G_B^{-1}$ [@MaiScherer2017]. We adopt the weaker
statements for simplicity and because our applications are solely under the case
of continuous, strictly monotonic margins. But see @Anderson2018 and 
@Popovic2018 for applications of copulas outside this case.

@@

Normal and Frank copula families 
are known to have symmetric tail dependence,
i.e., strengths of dependence in the lower and upper tails are the same, for all copulas
in these families [@nelsen2006_copula]. C copulas are known to have stronger lower- than 
upper-tail dependence, and survival Clayton copulas, being rotations of C copulas by 
$180^\circ$, have stronger upper- than lower tail dependence [@nelsen2006_copula].
Using the `iRho` function from the `copula` package, for each of the four families Normal, 
Frank, Clayton, and survival Clayton, we obtained parameters for which the expected value of 
Spearman's $\rho$ was, separately, $0,0.1,0.2,...,0.9$. For each copula family and for each 
of the selected parameter values we generated  $n$ points, 100 times, and thereby 
computed 100 values of the 
statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$. This exercise 
was repeated, separately, 
for $n=1000$ and for $n=35$, these cases being chosen, respectively, to reflect 
asymptotic behavior of the 
statistics and their behavior on the small datasets which are common in some fields
of ecology. The mean and standard error of the three difference statistics were computed for 
each copula family, parameter value, and value of $n$, and results were plotted against 
Spearman's rho (Figs \ref{fig_stat_testing35} and \ref{fig_stat_testing1000}), 
the expectation being that the difference statistics would be positive for 
C copulas, negative for survival Clayton copulas (except for the case of $\rho=0$, a 
boundary case of these families corresponding to independence), and not significantly different 
from $0$ for Frank and Normal 
copulas.
We used a $t$-test with a significance threshold of $0.05$ to judge significance here.
These expectations were fulfilled (Figs \ref{fig_stat_testing35} and 
\ref{fig_stat_testing1000}).

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/stat_results/stat_testing/res_pt35_rankT.pdf}
\caption[Testing nonparametric statistics with 35 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence of the Normal and Frank copulas. See Appendix  \ref{Test_npa_stat} for details. $n=35$. \label{fig_stat_testing35}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/stat_results/stat_testing/res_pt1000_rankT.pdf}
\caption[Testing nonparametric statistics with 1000 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence of the Normal and Frank copulas. See Appendix  \ref{Test_npa_stat} for details. $n=1000$.\label{fig_stat_testing1000}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/BMR_results/birds/birds_numbin_10_bivfunctionplot.pdf}
\caption[Nonparametric tests against an N-copula null, bird body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for bird body mass and BMR data. Same format as Fig. \ref{MT-fig_soilCN_nonparam}, see caption of that figure. \label{fig_bmr_birds_nonparam}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/BMR_results/mammals/mammals_numbin_10_bivfunctionplot.pdf}
\caption[Nonparametric tests against an N-copula null, mammal body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for mammal body mass and BMR data. Same format as Fig. \ref{MT-fig_soilCN_nonparam}, see caption of that figure. \label{fig_bmr_mammals_nonparam}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Ceder_creek_results/2000cc_numbin_4_bivfunctionplot.pdf}
\caption[Nonparametric tests against an N-copula null, Cedar Creek data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for Cedar Creek data, year 2000. As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving Normal surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.25),\ldots,(0.75,1)$. If the red cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than. For instance, a value $>N$ (respectively, $<N$) means the value of the statistic on real data was greater than (respectively, less than) its value on $N$ surrogates. When the statistic was greater than 975 or less than 975 surrogate values, it indicates significance ($95\%$ confidence level). When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, weaker than) expected from a normal-copula null hypothesis.\label{fig_cc2000_nonparam}}
\end{center}
\end{figure}

@@

knitr::kable(tab_3biv_npa_summary, booktabs = T, escape=F,
             format = "latex",linesep = "", row.names=F,
      caption.short = "Nonparametric asymmetry tests against an Normal-copula null model with birds, mammals and Cedar Creek data",
      caption = "Nonparametric asymmetry tests against an Normal-copula null model with birds, mammals and Cedar creek data: for birds and mammals data, results should be interpreted as of soil C and N results (see Table \\ref{MT-tab_soilCN_asym} which shows stronger upper tail dependence). For Cedar Creek data, first slice ($FS$) and last slice ($LS$) stand for $(0,0.25)$ and $(0.75,1)$ range of lower and upper bounds. Though the non-parametric result does not indicate significantly stronger lower tail dependence for Cedar Creek data, the 
greater than (>) sign before the numbers do show the opposite direction in tail-dependence compared to other bivariate datasets.\\label{tab_3biv_npa_summary}")
```

@@

The values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for the soil
C and N data
were compared to distributions of their values on 1000 Kendall- or Spearman-preserving Normal surrogates
of the data,
separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$ 
(Fig. \ref{MT-fig_soilCN_nonparam}). Results show that tail dependence of data is stronger in 
both the lower and upper tails than would be expected under a null hypothesis of a normal copula.
The values of the statistics $\cor_{0,0.1}-\cor_{0.9,1}$, $\Ps_{0,0.1}-\Ps_{0.9,1}$ and
$\Dsq_{0.9,1}-\Dsq_{0,0.1}$ for the soil C and N data were also compared to distributions of their values
on 1000 Kendall- or Spearman-preserving Normal surrogates, separately in two 
comparisons (Table \ref{MT-tab_soilCN_asym}). Results showed that
upper-tail dependence was significantly stronger than lower-tail dependence.

For first three bivariate datasets in Table \ref{MT-tab_bivar_summary},
the statistic $P_{0,0.1}-P_{0.9,1}$ was computed for real data and 
for 1000 Kendall-preserving surrogate datasets with Normal copulas. Due to relatively smaller sample size,
for Cedar Creek data we used $P_{0,0.25}-P_{0.75,1}$.
For each dataset, the rank of the statistic for real-data in 
the distribution of surrogate values was computed. 
If ranks were $>975$ or $<25$, asymmetry of tail 
dependence was significant compared to an N-copula null hypothesis. Ranks were `r unname(tab_bivar_summary[10,])`
for, respectively, the soil C and N dataset, the bird body masses and metabolic rates dataset, 
the mammal body masses and metabolic rates dataset, and the Cedar Creek data for the year 2000. 
Thus the first three datasets showed stronger 
upper- than lower-tail dependence, and the Cedar Creek data showed the reverse 
(though not significantly), confirming the 
results based on model selection presented in the main text (see 
Table \ref{MT-tab_bivar_summary}, row 9). Model selection results for Cedar Creek showed 
only slightly stronger lower- than upper-tail dependence, consistent with the non-significant 
tendency toward lower-tail dependence also revealed by the non-parametric analysis for that
dataset.

@@

<!--table for AIC-weights of Normal copula used for fitting, green spruce aphid abundance data-->
```{r tab_count_AICw_normal, echo=F, results='markup'}
source("normal_cop_AICw.R")
knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_count),
             format='latex',linesep = "",digits=3,
             caption.short = "AIC weights, Normal copulas, green spruce aphid abundance data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the Normal copula family for each location pair for green spruce aphid abundance data, for comparison with Table \\ref{tab_count_AICw}. \\label{tab_count_AICw_normal}",
             booktabs=TRUE
             )
```

@@

<!--table for AIC-weights of Normal copula used for fitting, leaf plum curling aphid first flight data-->
```{r tab_ff_AICw_normal, echo=F, results='markup'}
source("normal_cop_AICw.R")
knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_ff),
             format='latex',linesep = "",digits=3,
             caption.short="AIC weights, Normal copulas, leaf-curling plum aphid first flight data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for leaf-curlung plum aphid first flight data, for comparison with Table \\ref{tab_ff_AICw}.\\label{tab_ff_AICw_normal}",
             booktabs=TRUE
             )
```

@@

<!--table for AIC-weights of Normal copula used for fitting, Ceratium furca abundance data-->
```{r tab_pns_AICw_normal,echo=F,results="markup"}
source("normal_cop_AICw.R")
h<-normal_cop_AICw_matrix(r=RES_plankton_north_sea)
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="AIC weights, Normal copulas, \\textit{Ceratium furca} abundance data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the Normal copula family for each location pair for \\textit{Ceratium furca} abundance data, for comparison with Table \\ref{tab_pns_AICw}.\\label{tab_pns_AICw_normal}",
             booktabs=TRUE)
```

@@

<!--table for AIC-weights of Normal copula used for fitting, methane-flux data-->
```{r tab_methane_AICw_normal,echo=F,results="markup"}
source("normal_cop_AICw.R")
h<-normal_cop_AICw_matrix(r=RES_methane)
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="AIC weights, Normal copulas, methane-flux data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the Normal copula family for each location pair for methane-flux data, for comparison with Table \\ref{tab_methane_AICw}.\\label{tab_methane_AICw_normal}",
             booktabs=TRUE)
```

@@

The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
widely used spline correlogram method; the same scheme was also used by @walteretal2017 
as part of a method for 
performing model selection among matrix regression models. For each pair of 
locations $i \neq j$, the statistic
of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, 
$\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The 
values $s_{ij}$ were then arranged in
a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterizes the 
average strength of 
dependence, across all pairs of distinct locations, for data in the portion of the distributions 
given by the bounds $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, 
the mean characterizes average
asymmetry of tail dependence across all pairs of locations. 
Confidence intervals for the value of $\mean(S)$ were then computed as quantiles of a 
distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
revealed by the confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
N-copula null hypothesis.
Bootstrapping was performed by resampling the locations of measurement, with replacement. 
This corresponds to 
resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
[@bjornstadfalck2001 ; @walteretal2017].

@@

\noindent The matrices from which root character states and changes 
across phylogeny branches were randomly chosen were generated as follows.
For extreme right-tail dependent noise, we first produced
a 1,000,000 by $2$ matrix, $m$, filled with 2,000,000 
independent standard-normal random draws. Then, each row of $m$ 
was identified for a first set of modifications randomly 
and independently with a $50\%$ chance; 
for those rows identified, every element of the row was replaced with the 
absolute value of the first element of the row (including the first element itself).
For other rows, every element in the row was replaced by the negative of its
own absolute value. This produces right-tail dependent noise, normally distributed
for each of the two locations. 
If left-tail dependent noise was desired, we took the negative of the
entire matrix. The moderately left-tail dependent matrix was based on a Clayton copula 
with parameter chosen (using the `iRho` function in the `copula` package) so that
the Spearman correlation would be the same as for the extreme tail dependence 
datasets. The moderately right-tail dependent matrix and the symmetric tail dependence
matrix were constructed similarly, but using an survival Clayton copula and an Normal copula,
respectively.

@@

<!--Following chunk examine the consequence of non-Normal copula atructure on skewness using green spruce aphid abundance data-->
```{r skewness_count_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}

@@

<!--Following chunk examine the consequence of non-Normal copula atructure on skewness 
using leaf plum curling aphids' first flight data-->
```{r skewness_ff_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}

@@

<!--Following chunk examine the consequence of non-Normal copula atructure on skewness 
using ceratium furca's abundance data-->
```{r skewness_pns_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}

@@

<!--Following 4 chunks are examining consequence of non-Normal copula structure on Taylors law using made up data-->

>>>>

The data of Fig. \ref{MT-fig_cop_pedag1}A-B were generated to have standard normal marginals.
Fig. \ref{fig_transf_pearson_spearman}A can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
of both variables by the cummulative distribution function (cdf) of the standard normal distribution, and then by the inverse of 
the cdf of the gamma distribution with shape and scale parameters 2. 
The functions `pnorm` and `qgamma` in R were used.
Fig. \ref{fig_transf_pearson_spearman}B can be obtained from Fig. \ref{MT-fig_cop_pedag1}B in the same way.
The Pearson correlations differ substantially after transformation, 
but the Spearman correlations are unchanged (Figs \ref{MT-fig_cop_pedag1}A-B, \ref{fig_transf_pearson_spearman}).
A scatterplot very simlar to Fig. \ref{MT-fig_cop_pedag1}C can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
of both variables by the cdf of the standard normal distribution.
Likewise a scatterplot similar to Fig. \ref{MT-fig_cop_pedag1}D can be obtained from Fig. \ref{MT-fig_cop_pedag1}B in the same way.

@@

Now let $D$ be any bivariate copula and let 
$G_A$ and $G_B$ be univariate distribution functions, so that 
$D(G_A(x),G_B(y))$ is guaranteed by Sklar's theorem to be a distribution function.
We here show that if $G_A$ and $G_B$ are continuous and strictly monotonic, then
$D(G_A(x),G_B(y))$ is the distribution function of the random variable 
$(G_A^{-1}(U),G_B^{-1}(V))$, where $(U,V)$ now denotes the random
vector with uniform marginals associated with $D$ and
$G_A^{-1}$ and $G_B^{-1}$ are the inverses of $G_A$ and $G_B$.
The inverses exist because of the strict monotonicity assumption.
To get to the desired result, note that the distribution function of 
$(G_A^{-1}(U),G_B^{-1}(V))$ is $P[G_A^{-1}(U) \leq x, G_B^{-1}(V) \leq y]$,
which equals $P[U \leq G_A(x), V \leq G_B(y)]$, by strict monotonicity.
But this is $D(G_A(x),G_B(y))$ because $D$ is the distribution function of 
$(U,V)$. In fact, a stronger statement assuming
neither continuity nor strict monotonicity is possible, if one uses generalized
inverses for $G_A^{-1}$ and $G_B^{-1}$ [@MaiScherer2017]. We adopt the weaker
statements for simplicity and because our applications are solely under the case
of continuous, strictly monotonic margins. But see @Anderson2018 and 
@Popovic2018 for applications of copulas outside this case.

@@

Normal and Frank copula families 
are known to have symmetric tail dependence,
i.e., strengths of dependence in the lower and upper tails are the same, for all copulas
in these families [@nelsen2006_copula]. Clayton copulas are known to have stronger lower- than 
upper-tail dependence, and survival Clayton copulas, being rotations of Clayton copulas by 
$180^\circ$, have stronger upper- than lower tail dependence [@nelsen2006_copula].
Using the `iRho` function from the `copula` package, for each of the four families normal, 
Frank, Clayton, and survival Clayton, we obtained parameters for which the expected value of 
Spearman's $\rho$ was, separately, $0,0.1,0.2,...,0.9$. For each copula family and for each 
of the selected parameter values we generated  $n$ points, 100 times, and thereby 
computed 100 values of the 
statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$. This exercise 
was repeated, separately, 
for $n=1000$ and for $n=35$, these cases being chosen, respectively, to reflect 
asymptotic behavior of the 
statistics and their behavior on the small datasets which are common in some fields
of ecology. The mean and standard error of the three difference statistics were computed for 
each copula family, parameter value, and value of $n$, and results were plotted against 
Spearman's rho (Figs \ref{fig_stat_testing35} and \ref{fig_stat_testing1000}), 
the expectation being that the difference statistics would be positive for 
Clayton copulas, negative for survival Clayton copulas (except for the case of $\rho=0$, a 
boundary case of these families corresponding to independence), and not significantly different 
from $0$ for Frank and normal 
copulas.
We used a $t$-test with a significance threshold of $0.05$ to judge significance here.
These expectations were fulfilled (Figs \ref{fig_stat_testing35} and 
\ref{fig_stat_testing1000}).

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/stat_results/stat_testing/res_pt35_rankT.pdf}
\caption[Testing nonparametric statistics with 35 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence of the normal and Frank copulas. See Appendix  \ref{Test_npa_stat} for details. $n=35$. \label{fig_stat_testing35}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/stat_results/stat_testing/res_pt1000_rankT.pdf}
\caption[Testing nonparametric statistics with 1000 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence of the normal and Frank copulas. See Appendix  \ref{Test_npa_stat} for details. $n=1000$.\label{fig_stat_testing1000}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/BMR_results/birds/birds_numbin_10_bivfunctionplot.pdf}
\caption[Nonparametric tests against a normal copula null, bird body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for bird body mass and BMR data. Same format as Fig. \ref{MT-fig_soilCN_nonparam}, see caption of that figure. \label{fig_bmr_birds_nonparam}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/BMR_results/mammals/mammals_numbin_10_bivfunctionplot.pdf}
\caption[Nonparametric tests against a normal copula null, mammal body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for mammal body mass and BMR data. Same format as Fig. \ref{MT-fig_soilCN_nonparam}, see caption of that figure. \label{fig_bmr_mammals_nonparam}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Ceder_creek_results/2000cc_numbin_4_bivfunctionplot.pdf}
\caption[Nonparametric tests against a normal copula null, Cedar Creek data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for Cedar Creek data, year 2000. As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.25),\ldots,(0.75,1)$. If the red cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than. For instance, a value $>N$ (respectively, $<N$) means the value of the statistic on real data was greater than (respectively, less than) its value on $N$ surrogates. When the statistic was greater than 975 or less than 975 surrogate values, it indicates significance ($95\%$ confidence level). When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, weaker than) expected from a normal-copula null hypothesis.\label{fig_cc2000_nonparam}}
\end{center}
\end{figure}

@@

knitr::kable(tab_3biv_npa_summary, booktabs = T, escape=F,
             format = "latex",linesep = "", row.names=F,
      caption.short = "Nonparametric asymmetry tests against a normal copula null model with birds, mammals and Cedar Creek data",
      caption = "Nonparametric asymmetry tests against an normal copula null model with birds, mammals and Cedar creek data: for birds and mammals data, results should be interpreted as of soil C and N results (see Table \\ref{MT-tab_soilCN_asym} which shows stronger upper tail dependence). For Cedar Creek data, first slice ($FS$) and last slice ($LS$) stand for $(0,0.25)$ and $(0.75,1)$ range of lower and upper bounds. Though the non-parametric result does not indicate significantly stronger lower tail dependence for Cedar Creek data, the 
greater than (>) sign before the numbers do show the opposite direction in tail-dependence compared to other bivariate datasets.\\label{tab_3biv_npa_summary}")
```

@@

The values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for the soil
C and N data
were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates
of the data,
separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$ 
(Fig. \ref{MT-fig_soilCN_nonparam}). Results show that tail dependence of data is stronger in 
both the lower and upper tails than would be expected under a null hypothesis of a normal copula.
The values of the statistics $\cor_{0,0.1}-\cor_{0.9,1}$, $\Ps_{0,0.1}-\Ps_{0.9,1}$ and
$\Dsq_{0.9,1}-\Dsq_{0,0.1}$ for the soil C and N data were also compared to distributions of their values
on 1000 Kendall- or Spearman-preserving normal surrogates, separately in two 
comparisons (Table \ref{MT-tab_soilCN_asym}). Results showed that
upper-tail dependence was significantly stronger than lower-tail dependence.

For first three bivariate datasets in Table \ref{MT-tab_bivar_summary},
the statistic $P_{0,0.1}-P_{0.9,1}$ was computed for real data and 
for 1000 Kendall-preserving surrogate datasets with normal copulas. Due to relatively smaller sample size,
for Cedar Creek data we used $P_{0,0.25}-P_{0.75,1}$.
For each dataset, the rank of the statistic for real-data in 
the distribution of surrogate values was computed. 
If ranks were $>975$ or $<25$, asymmetry of tail 
dependence was significant compared to a normal copula null hypothesis. Ranks were `r unname(tab_bivar_summary[10,])`
for, respectively, the soil C and N dataset, the bird body masses and metabolic rates dataset, 
the mammal body masses and metabolic rates dataset, and the Cedar Creek data for the year 2000. 
Thus the first three datasets showed stronger 
upper- than lower-tail dependence, and the Cedar Creek data showed the reverse 
(though not significantly), confirming the 
results based on model selection presented in the main text (see 
Table \ref{MT-tab_bivar_summary}, row 9). Model selection results for Cedar Creek showed 
only slightly stronger lower- than upper-tail dependence, consistent with the non-significant 
tendency toward lower-tail dependence also revealed by the non-parametric analysis for that
dataset.

@@

<!--table for AIC-weights of normal copula used for fitting, green spruce aphid abundance data-->
```{r tab_count_AICw_normal, echo=F, results='markup'}
source("normal_cop_AICw.R")
knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_count),
             format='latex',linesep = "",digits=3,
             caption.short = "AIC weights, normal copulas, green spruce aphid abundance data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for green spruce aphid abundance data, for comparison with Table \\ref{tab_count_AICw}. \\label{tab_count_AICw_normal}",
             booktabs=TRUE
             )
```

@@

<!--table for AIC-weights of normal copula used for fitting, leaf plum curling aphid first flight data-->
```{r tab_ff_AICw_normal, echo=F, results='markup'}
source("normal_cop_AICw.R")
knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_ff),
             format='latex',linesep = "",digits=3,
             caption.short="AIC weights, normal copulas, leaf-curling plum aphid first flight data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for leaf-curlung plum aphid first flight data, for comparison with Table \\ref{tab_ff_AICw}.\\label{tab_ff_AICw_normal}",
             booktabs=TRUE
             )
```

@@

<!--table for AIC-weights of normal copula used for fitting, Ceratium furca abundance data-->
```{r tab_pns_AICw_normal,echo=F,results="markup"}
source("normal_cop_AICw.R")
h<-normal_cop_AICw_matrix(r=RES_plankton_north_sea)
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="AIC weights, normal copulas, \\textit{Ceratium furca} abundance data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for \\textit{Ceratium furca} abundance data, for comparison with Table \\ref{tab_pns_AICw}.\\label{tab_pns_AICw_normal}",
             booktabs=TRUE)
```

@@

<!--table for AIC-weights of normal copula used for fitting, methane-flux data-->
```{r tab_methane_AICw_normal,echo=F,results="markup"}
source("normal_cop_AICw.R")
h<-normal_cop_AICw_matrix(r=RES_methane)
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="AIC weights, normal copulas, methane-flux data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for methane-flux data, for comparison with Table \\ref{tab_methane_AICw}.\\label{tab_methane_AICw_normal}",
             booktabs=TRUE)
```

@@

The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
widely used spline correlogram method; the same scheme was also used by @walteretal2017 
as part of a method for 
performing model selection among matrix regression models. For each pair of 
locations $i \neq j$, the statistic
of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, 
$\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The 
values $s_{ij}$ were then arranged in
a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterizes the 
average strength of 
dependence, across all pairs of distinct locations, for data in the portion of the distributions 
given by the bounds $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, 
the mean characterizes average
asymmetry of tail dependence across all pairs of locations. 
Confidence intervals for the value of $\mean(S)$ were then computed as quantiles of a 
distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
revealed by the confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
normal copula null hypothesis.
Bootstrapping was performed by resampling the locations of measurement, with replacement. 
This corresponds to 
resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
[@bjornstadfalck2001 ; @walteretal2017].

@@

\noindent The matrices from which root character states and changes 
across phylogeny branches were randomly chosen were generated as follows.
For extreme right-tail dependent noise, we first produced
a 1,000,000 by $2$ matrix, $m$, filled with 2,000,000 
independent standard-normal random draws. Then, each row of $m$ 
was identified for a first set of modifications randomly 
and independently with a $50\%$ chance; 
for those rows identified, every element of the row was replaced with the 
absolute value of the first element of the row (including the first element itself).
For other rows, every element in the row was replaced by the negative of its
own absolute value. This produces right-tail dependent noise, normally distributed
for each of the two locations. 
If left-tail dependent noise was desired, we took the negative of the
entire matrix. The moderately left-tail dependent matrix was based on a Clayton copula 
with parameter chosen (using the `iRho` function in the `copula` package) so that
the Spearman correlation would be the same as for the extreme tail dependence 
datasets. The moderately right-tail dependent matrix and the symmetric tail dependence
matrix were constructed similarly, but using an survival Clayton copula and an normal copula,
respectively.

@@

<!--Following chunk examine the consequence of non-normal copula atructure on skewness using green spruce aphid abundance data-->
```{r skewness_count_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}

@@

<!--Following chunk examine the consequence of non-normal copula atructure on skewness 
using leaf plum curling aphids' first flight data-->
```{r skewness_ff_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}

@@

<!--Following chunk examine the consequence of non-normal copula atructure on skewness 
using ceratium furca's abundance data-->
```{r skewness_pns_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}

@@

<!--Following 4 chunks are examining consequence of non-normal copula structure on Taylors law using made up data-->
